---
title: 'A hidden deadly venomous insect: first eco-epidemiological assessment and
  risk mapping of lonomism in Argentina'
author: "MG Casafús, MM Favalesso, MA Gritti, JM Coronel, ATB Guimarães and ME Peichoto"
output:
  html_document:
    df_print: paged
---

## R packages

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
pacman::p_load(readxl, ggplot2, spaMM, lme4, MASS, sjPlot, pscl, corrplot, Hmisc, PerformanceAnalytics, mctest, fitdistrplus, GGally, ggExtra, sf, viridis, cartography, sp, ggspatial, binr, gganimate, plotly, gifski, magick, rgdal, rgeos, Rcpp, raster, ncdf4, RNetCDF, climateR, AOI, kableExtra, scico, tibble, cowplot, ggsn, reshape2, tidyverse, spatstat, maptools, KernSmooth, GISTool, dichromat, crayon)
```

## ***Table 1***: Socio-demographic profile of victims of lonomic accidents in Misiones, Argentina.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
## --- data analysis --- ##
# Sex
s = c(11, 29)

sex = data.frame(Variable = c("Sex", ""),
                 Category = c("F", "M"),
                 n = s,
                 Residuals = c(chisq.test(s)$residuals %>% round(2)),
                 X2 = c(chisq.test(s)$statistic %>% round(2), paste0("(df = ", chisq.test(s)$parameter, ")")),
                 p =  c(chisq.test(s)$p.value %>% round(4), "")
                 )

# Age group
a = c(15, 11, 2, 4, 2, 4, 1)

age = data.frame(Variable = c("Age group", rep("", 6)), 
                 Category = c("0-10", "11-20", "21-30", "31-40", "41-50", "51-60", "61-70"),
                 n = a, 
                 Residuals = c(chisq.test(a)$residuals %>% round(2)), 
                 X2 = c(chisq.test(a)$statistic %>% round(2), paste0("(df = ", chisq.test(a)$parameter, ")"), rep("",5)), 
                 p = c(chisq.test(a)$p.value %>% round(4), rep("",6))
                 )

# Area
ar = c(6, 17, 13)

area = data.frame(Variable = c("Area", rep("", 2)), 
                 Category = c("Peri-Urban", "Rural", "Forest"),
                 n = ar, 
                 Residuals = c(chisq.test(ar)$residuals %>% round(2)), 
                 X2 = c(chisq.test(ar)$statistic %>% round(2), paste0("(df = ", chisq.test(ar)$parameter, ")"), ""), 
                 p = c(chisq.test(ar)$p.value %>% round(4), rep("",2))
)

# Departament
d = c(1, 1, 8, 6, 10, 2, 8)

departament = data.frame(Variable = c("Departament", rep("", 6)), 
                  Category = c("25 de Mayo", "Alem", "Cainguás", "Gral. Manuel Belgrano", "Guaraní", "Oberá", "San Pedro"),
                  n = d, 
                  Residuals = c(chisq.test(d)$residuals %>% round(2)), 
                  X2 = c(chisq.test(d)$statistic %>% round(2), paste0("(df = ", chisq.test(d)$parameter, ")"), rep("", 5)), 
                  p = c(chisq.test(d)$p.value %>% round(4), rep("", 6))
)

# Circumstances
c = c(12, 23)

circumstances = data.frame(Variable = c("Circumstances", ""), 
                  Category = c("During work activity", "During recreational activity"),
                  n = c, 
                  Residuals = c(chisq.test(c)$residuals %>% round(2)), 
                  X2 = c(chisq.test(c)$statistic %>% round(2), paste0("(df = ", chisq.test(c)$parameter, ")")), 
                  p = c(chisq.test(c)$p.value %>% round(4), "")
)

# Time of the day
t = c(12, 22)

time = data.frame(Variable = c("Time of the day", ""), 
                  Category = c("06:00 - 12:00h", "12:01 - 18:00h"),
                  n = t, 
                  Residuals = chisq.test(t)$residuals %>% round(2), 
                  X2 = c(chisq.test(t)$statistic %>% round(2), paste0("(df = ", chisq.test(t)$parameter, ")")), 
                  p = c(chisq.test(t)$p.value %>% round(4), "")
)

# Affected body part
af = c(8, 31)

affected = data.frame(Variable = c("Affected body part", ""), 
                  Category = c("Lower limb", "Upper limb"),
                  n = af, 
                  Residuals = chisq.test(af)$residuals %>% round(2), 
                  X2 = c(chisq.test(af)$statistic %>% round(2), paste0("(df = ", chisq.test(af)$parameter, ")")), 
                  p = c(chisq.test(af)$p.value %>% round(4), "")
)

## --- Table 1: Results --- ##
tab1 = rbind(sex, age, area, departament, circumstances, time, affected)

rownames(tab1) = NULL

tab1 %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## ***Table 2*****:** Median and interquartile interval of abiotic variables extracted from georeferenced points of occurrence/accident cases of *Lonomia* spp. in Misiones, Argentina.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
## --- data analysis --- ##

## Shape of Misiones - Argentina 
## downloaded from:
## https://www.ign.gob.ar/NuestrasActividades/InformacionGeoespacial/CapasSIG

dep = st_read("departamento/departamento.shp")
prov = st_read("provincia/provincia.shp")
mis = prov[prov$nam == "Misiones",]

dep2 = st_crop(dep, mis)
misiones = dep2[dep2$fdc == "Ministerio de Ecología",]

## Occurence data
pontos = read_excel("amostragem.xlsx")
names(pontos) = c('id', 'life fase', 'x', 'y', 'month', 'year')

## --- TerraClimate data --- ## 
var = getTerraClim(AOI = misiones, param = c("tmax", "tmin", "prcp", "aet", "srad"),
                   startDate = "2014-01-01", endDate   = "2019-12-31")

## --- Extract the values for the points --- ##
l = list()

for(i in names(var)){
  for(j in c(1:dim(pontos)[1])){
    am = 
      if(pontos[j, 'month'] < 10){paste0("X", pontos[j, "year"], ".0", pontos[j, 'month'])
      }else{paste0("X", pontos[j, "year"], ".", pontos[j, 'month'])}
    
    l[[i]][[j]] = raster::extract(var[[i]][[am]], pontos[j, c(3,4)])
  }
}


## --- Table 2: Results --- ##
tab2 = data.frame(Sample = c("Maximum Temperature (ºC)", "Minimum Temperature (ºC)", "Precipitation (mm)", "Evapotranspiration (mm/day)", "Radiation (Wm^-2)"),
                  Median = c(round(quantile(l$tmax %>% unlist)[[3]],1),
                             round(quantile(l$tmin %>% unlist)[[3]],1),
                             round(quantile(l$prcp %>% unlist)[[3]],1),
                             round(quantile(l$aet %>% unlist)[[3]],1),
                             round(quantile(l$srad %>% unlist)[[3]],1)),
                  Q1 = c(round(quantile(l$tmax %>% unlist)[[2]],1),
                         round(quantile(l$tmin %>% unlist)[[2]],1),
                         round(quantile(l$prcp %>% unlist)[[2]],1),
                         round(quantile(l$aet %>% unlist)[[2]],1),
                         round(quantile(l$srad %>% unlist)[[2]],1)),
                  Q3 = c(round(quantile(l$tmax %>% unlist)[[4]],1),
                         round(quantile(l$tmin %>% unlist)[[4]],1),
                         round(quantile(l$prcp %>% unlist)[[4]],1),
                         round(quantile(l$aet %>% unlist)[[4]],1),
                         round(quantile(l$srad %>% unlist)[[4]],1))
)

tab2 %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
tab2 = data.frame(Sample = c("Maximum Temperature (ºC)", "Minimum Temperature (ºC)", "Precipitation (mm)", "Evapotranspiration (mm/day)", "Radiation (Wm^-2)"),
                 Median = c(28.4, 15.8, 148.0, 115.0, 206.0),
                 Q1 = c(27.0, 14.7, 117.0, 90.8, 173.3),
                 Q3 = c(29.6, 17.3, 210.8, 132.0, 237.3))

tab2 %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## ***Figure 3***: Accumulated occurrences of *Lonomia* spp. in the months of the study period (January 2014 to May 2020). Rasterized demonstration of evapotranspiration, rainfall precipitation, solar radiation, maximum and minimum temperatures in periods of high level of occurrences (January to May; September to December) and absence of occurrence (June to August).

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
## --- data analysis --- ##
n = list()
a = 0

for(i in c("02", "07", "12")){
  for(j in c(2014:2019)){
    n[[a + 1]] = paste0("X", j, ".", i)
    a = a + 1
  }
}

m = n %>% unlist()


## Min. Temp. (ºC) --
rtmin2 = var$tmin[[m[1:6]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  tibble::as_tibble() %>% 
  ggplot2::ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = "vik", direction = 1, limit = c(6, 22)) +
  coord_sf() + theme_bw() + labs(x = NULL, y = NULL, fill = "", title = "Min. Temp. (ºC) - February") 

rtmin7 = var$tmin[[m[7:12]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = "vik", direction = 1, limit = c(6, 22)) +
  coord_sf() + theme_bw() + labs(x = NULL, y = NULL, fill = "", title = "Min. Temp. (ºC) - July")

rtmin12 = var$tmin[[m[13:18]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = "vik", direction = 1, limit = c(6, 22)) +
  coord_sf() + theme_bw() + labs(x = NULL, y = NULL, fill = "", title = "Min. Temp. (ºC) - December")


## Max. Temp. (ºC) --

rtmax2 = var$tmax[[m[1:6]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[27], direction = 1, limit = c(18.5, 33.5)) +
  coord_sf() + theme_bw() + labs(x = NULL, y = NULL, fill = "", title = "Max. Temp. (ºC) - February") 

rtmax7 = var$tmax[[m[7:12]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[27], direction = 1, limit = c(18.5, 33.5)) +
  coord_sf() + theme_bw() + labs(x = NULL, y = NULL, fill = "", title = "Max. Temp. (ºC) - July")

rtmax12 = var$tmax[[m[13:18]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[27], direction = 1, limit = c(18.5, 33.5)) +
  coord_sf() + theme_bw() + labs(x = NULL, y = NULL, fill = "", title = "Max. Temp. (ºC) - December")

## Radiation (Wm^-2) --
rrad2 = var$srad[[m[1:6]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[16], direction = 1, limit = c(99, 265)) +
  coord_sf() + theme_bw() + 
  labs(x = NULL, y = NULL, fill = "", title = "Radiation (Wm^-2) - February") 

rrad7 = var$srad[[m[7:12]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[16], direction = 1, limit = c(99, 265)) +
  coord_sf() + theme_bw() + 
  labs(x = NULL, y = NULL, fill = "", title = "Radiation (Wm^-2) - July")

rrad12 = var$srad[[m[13:18]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[16], direction = 1, limit = c(99, 265)) +
  coord_sf() + theme_bw() + 
  labs(x = NULL, y = NULL, fill = "", title = "Radiation (Wm^-2) - December")


## Precipitation (mm) --

rprec2 = var$prcp[[m[1:6]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[21], direction = -1, limit = c(56, 240)) +
  coord_sf() + theme_bw() + 
  labs(x = NULL, y = NULL, fill = "", title = "Precipitation (mm) - February") 

rprec7 = var$prcp[[m[7:12]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[21], direction = -1, limit = c(56, 240)) +
  coord_sf() + theme_bw() + 
  labs(x = NULL, y = NULL, fill = "", title = "Precipitation (mm) - July")

rprec12 = var$prcp[[m[13:18]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[21], direction = -1, limit = c(56, 240)) +
  coord_sf() + theme_bw() + 
  labs(x = NULL, y = NULL, fill = "", title = "Precipitation (mm)- December")

## Evapotrans. (mm) --

rev2 = var$aet[[m[1:6]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[9], direction = -1, limit = c(44, 155)) +
  coord_sf() + theme_bw() + 
  labs(x = NULL, y = NULL, fill = "", title = "Evapotrans. (mm) - February") 

rev7 = var$aet[[m[7:12]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[9], direction = -1, limit = c(44, 155)) +
  coord_sf() + theme_bw() + 
  labs(x = NULL, y = NULL, fill = "", title = "Evapotrans. (mm) - July")

rev12 = var$aet[[m[13:18]]] %>% 
  stack() %>% 
  raster::crop(misiones) %>% 
  raster::mask(misiones) %>% 
  stackApply(indices = rep(1, length(m)), fun = median, na.rm = T) %>%
  rasterToPoints() %>%
  as_tibble() %>% 
  ggplot() + geom_raster(aes(x = x, y = y, fill = index_1)) +
  scale_fill_scico(palette = scico_palette_names()[9], direction = -1, limit = c(44, 155)) +
  coord_sf() + theme_bw() + 
  labs(x = NULL, y = NULL, fill = "", title = "Evapotrans. (mm) - December")

## --- Saving --- ##
ggsave("rtmin2.tiff", rtmin2, he = 15, wi = 20, un = "cm", dpi = 300)
ggsave("rtmin7.tiff", rtmin7, he = 15, wi = 20, un = "cm", dpi = 300)
ggsave("rtmin12.tiff", rtmin12, he = 15, wi = 20, un = "cm", dpi = 300)

ggsave("rtmax2.tiff", rtmax2, he = 15, wi = 20, un = "cm", dpi = 300)
ggsave("rtmax7.tiff", rtmax7, he = 15, wi = 20, un = "cm", dpi = 300)
ggsave("rtmax12.tiff", rtmax12, he = 15, wi = 20, un = "cm", dpi = 300)

ggsave("rrad2.tiff", rrad2, he = 15, wi = 20, un = "cm", dpi = 300)
ggsave("rrad7.tiff", rrad7, he = 15, wi = 20, un = "cm", dpi = 300)
ggsave("rrad12.tiff", rrad12, he = 15, wi = 20, un = "cm", dpi = 300)

ggsave("rprec2.tiff", rprec2, he = 15, wi = 20, un = "cm", dpi = 300)
ggsave("rprec7.tiff", rprec7, he = 15, wi = 20, un = "cm", dpi = 300)
ggsave("rprec12.tiff", rprec12, he = 15, wi = 20, un = "cm", dpi = 300)

ggsave("rev2.tiff", rev2, he = 15, wi = 20, un = "cm", dpi = 300)
ggsave("rev7.tiff", rev7, he = 15, wi = 20, un = "cm", dpi = 300)
ggsave("rev12.tiff", rev12, he = 15, wi = 20, un = "cm", dpi = 300)
```

The bar graph was designed using [Microsoft Excel](https://www.microsoft.com/en-us/microsoft-365/excel) software.

The maps were related to the bar graph using the [Inkscape software](https://inkscape.org/pt-br/).

## ***Figure 4*****:** **A**- Distribution by area of accidental/occasional occurrence of *Lonomia* spp. in Misiones, Argentina. **B**- Distribution by area and specific space inside each area where lonomism cases happened in Misiones, Argentina. The superscripted letters 'a' and 'b' refer to the highest and lowest (respectively) classification of frequencies according to the adjusted residual post-hoc test.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
c(28, 24, 14, 3) %>% chisq.test()
chisq.test(c(28, 24, 14, 3))$residuals

```

Bar charts A and B were prepared using [Microsoft Excel software](Bar%20charts%20A%20and%20B%20were%20prepared%20using%20Microsoft%20Excel%20software.).

Graphics A and B were merged into a single image using the [Inkscape software](https://inkscape.org/pt-br/).

## ***Table 3***: Host plants (n=50) for *Lonomia* larvae in Misiones, Argentina.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
## --- data analysis --- ##
plant = read_excel("plant.xlsx") %>%
  as_tibble()

## Species --
spc = plant %>% filter(Variable == 'Species') %>% select(3)

spcd = data.frame(Variable = c("Species", rep("", 24)),
                 Category = plant %>% filter(Variable == 'Species')%>% select(2),
                 n = spc,
                 Residuals = c(chisq.test(spc)$residuals %>% round(2)),
                 X2 = c(chisq.test(spc)$statistic %>% round(3), paste0("(df = ", chisq.test(spc)$parameter, ")"), rep("", 23)),
                 p =  c(chisq.test(spc)$p.value %>% round(4), rep("", 24))
)

## Family --
fam = plant %>% filter(Variable == 'Family') %>% select(3)



fami = data.frame(Variable = c("Family", rep("", 15)),
                  Category = c(plant %>% filter(Variable == 'Family') %>% select(2)),
                  n = fam,
                  Residuals = c(chisq.test(fam)$residuals %>% round(2)),
                  X2 = c(chisq.test(fam)$statistic %>% round(3), paste0("(df = ", chisq.test(fam)$parameter, ")"), rep("", 14)),
                  p =  c(chisq.test(fam)$p.value %>% round(4), rep("", 15))
)

## Status --
st = plant %>% filter(Variable == 'Status') %>% select(3)

stat = data.frame(Variable = c("Status", ""),
                  Category = c(plant %>% filter(Variable == 'Status') %>% select(2)),
                  n = st,
                  Residuals = c(chisq.test(st)$residuals %>% round(2)),
                  X2 = c(chisq.test(st)$statistic %>% round(3), paste0("(df = ", chisq.test(st)$parameter, ")")),
                  p =  c(chisq.test(st)$p.value %>% round(4), "")
)

st = plant %>% filter(Variable == 'Status') %>% select(3)

stat = data.frame(Variable = c("Status", ""),
                  Category = c(plant %>% filter(Variable == 'Status') %>% select(2)),
                  n = st,
                  Residuals = c(chisq.test(st)$residuals %>% round(2)),
                  X2 = c(chisq.test(st)$statistic %>% round(3), paste0("(df = ", chisq.test(st)$parameter, ")")),
                  p =  c(chisq.test(st)$p.value %>% round(4), "")
)

## Tree type -- 
tree = plant %>% filter(Variable == "Tree type") %>% select(3)

tr = data.frame(Variable = c("Tree type", ""),
                Category = c(plant %>% filter(Variable == "Tree type") %>% select(2)),
                n = tree,
                Residuals = c(chisq.test(tree)$residuals %>% round(2)),
                X2 = c(chisq.test(tree)$statistic %>% round(3), paste0("(df = ", chisq.test(tree)$parameter, ")")),
                p =  c(chisq.test(tree)$p.value %>% round(4), "")
)

## --- Table 3: Results --- ## 
tab3 = rbind(spcd, fami, stat, tr)
rownames(tab3) = NULL

tab3 %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

## ***Figure 6***: Lonomism risk map for Misiones, Argentina (January 26th, 2014 - May 8th, 2020) using kernel density estimation.

```{r eval=FALSE, message=FALSE, warning=FALSE, include=TRUE}
## Data --
## Misiones
misiones2 = as_Spatial(mis)
citys_misiones = as_Spatial(misiones)

## Data points 
pnts = read_excel("pnts.xlsx")
coord = data.frame(pnts[,11], pnts[,10])

## Kernel denity estimate map --- 
## Density map
est = bkde2D(coord, 
             bandwidth = c(0.15, 0.15), 
             gridsize = c(1000L, 1000L)) 

## Transforming the density map into a raster
est.raster = raster(list(x=est$x1, y=est$x2, z=est$fhat))

## We are pairing the projection of the map with misiones
projection(est.raster) = projection(misiones2) 

## Cutting from the size of Misiones
result_raster = raster::mask(crop(est.raster, misiones2), misiones2)

## --- Result --- ##
## Statistical map for the manuscript ---
col_map = colorRampPalette(c("grey95", "grey65", "grey50"))(10) # 

# Plotting kernel
plot(c(-56.5, -52), c(-28.2, -25.5), axes = F, ann = F, type = "n") 
box(col = "black") 
plot(misiones, lwd = 2, border = NA, col = "grey95", add = T) 
plot(result_raster, add = TRUE, legend = F, col = col_map) 

# The divisions
plot(citys_misiones, lwd = 1.8, border = 'grey30', add = T) 

# Points
# Form and color for the points
pnts_form = ifelse(pnts$Coord == "Coordinate", 22, 24)
pnts_color = ifelse(pnts$ID == "Moth", "#FFFF00", 
                    ifelse(pnts$ID == "Larva", "#00FF00", "#FF3333"))

points(coord, pch = pnts_form, 
       bg = pnts_color, col = "black", cex = 1.3, lwd = 0.9)

# Division number names
centro = coordinates(citys_misiones)
citys_misiones$x = centro[,1]
citys_misiones$y = centro[,2]

names_city = data.frame(citys_misiones)
names_city$nam[12] = "Venticinco de Mayo"
names_city$cod = c(9, 6, 16, 12, 3, 7, 4, 2, 11, 8, 13,
                   17, 14, 10, 15, 5, 1)

text(names_city[,'x']-0.02, names_city[,'y']-0.09,  names_city$cod,
     cex=0.8, pos=3, col="black", bg ='white', font = 2) 

# Legends: 
text(-53.17, -25.45, "LEGEND", cex = 0.8, font = 2)

legend(-53.5, -25.58, title = expression(bold('      Regions')), cex = 0.6, 
       bg = NA,
       box.col = NA, title.adj = 0, adj = 0, text.font = 1,
       legend = c(paste0(sort(names_city$cod), " - ", str_sort(names_city$nam))))

legend(-53.5, -27.1, pch = 21, inset = .02, 
       legend = c("Moth", "Larvae", "Accident"), 
       pt.bg = c("#FFFF00",  "#00FF00", "#FF3333"), bty = "topleft", bg = NA, 
       cex = 0.6, pt.cex = 1.2, pt.lwd = 0.5,
       box.col = NA, title = expression(bold("Sample type")))

legend(-53.5, -27.53, legend = c("Coordinate", "Centroid"), pt.cex = 1,
       pt.lwd = 0.7,
       pch = c(22, 24), bg = NA, pt.bg = "grey60", cex = 0.6, 
       box.col = NA, title = expression(bold("Sample precision")))

plot(result_raster, legend.only = T, col = col_map,
     legend.args = list(text = expression(bold('Density ')), 
                        side = 3, line = 0.1, cex = 0.61), 
     axis.args = list(cex.axis = 0.61, col = "black", labels = c("Low", "High"), 
                      at = c(1.875618e-05, 1.461), lwd = 0.1), bg = "white",
     smallplot = c(0.655, 0.8, 0.235, 0.249), horizon = T,
     labels = F)
```

## Responsible for data analysis

**Ana Tereza Bittencourt Guimarães**

`r icons::icon_style(icons::fontawesome("envelope", style = "solid"), fill = "#33CCCC")` [anatbguimaraes\@gmail.com](anatbguimaraes@gmail.com)

`r icons::icon_style(icons::fontawesome("researchgate"), fill = "#33CCCC")` [/Ana-Guimaraes-27](https://www.researchgate.net/profile/Ana-Guimaraes-27)

**Marília Melo Favalesso**

`r icons::icon_style(icons::fontawesome("envelope", style = "solid"), fill = "#33CCCC")` [mariliabioufpr\@gmail.com](mariliabioufpr@gmail.com)

`r icons::icon_style(icons::fontawesome("globe-americas", style = "solid"), fill = "#33CCCC")` [www.mmfava.com](www.mmfava.com)

`r icons::icon_style(icons::fontawesome("github"), fill = "#33CCCC")` [Github: mmfava](https://github.com/mmfava)
